Music Versus No Music Effectiveness On Cognitive Response Time and Typing Efficiency

### 2024-05-07 Latex Source/paper.tex
Update paper.tex

During the cognitive task component of the study, participants performed a Quantified Behavioral test or QBTest while their eye movements were tracked by the EyeWriter 2.0.
The QBTest is a computerized test designed to assess various aspects of cognitive function, including attention, impulsivity, and response control.
The user will then see Figure~\ref{QBTest}. Figure~\ref{QBTest}a is the screen that will be displayed before the participant is ready to begin the QBTest.
Figure~\ref{QBTest}b is what the user will see during the test. Figure~\ref{QBTest}b also displays what the user will see when they are correct and incorrect.
Figure~\ref{QBTest}c displays for only two minutes but shows the user their score.
% MAYBE ADD MORE ABOUT THE QBTEST?

\section{Results and Discussion}
Within-subjects testing was used for our ten participants. Individually the typing test and QBTest scores were not statistically significant, whereas 
% WHATEVER TYPE OF TESTING IS USED (ANOVA)
the ANOVA from the heatmap with the eyetracker, the results were statistically significant. %sounds clunky might change
The ANOVA for the typing test was done twice for the dependent variables, time and accuracy, as GoStats did not allow for multiple 
dependent variables. The accuracy for all participants were not necessarily an indicator of good results as all the participants were 
allowed to use the backspace to re-try any letters that were mis-typed.

@@

While not statistically significant, the \textit{p} value for time was 0.1172 whereas the \textit{p} value for accuracy was 0.2334.
The bar graph in Figure~\ref{timeL} for time in seconds for the typing test was quite telling. The condition for no music was consistently faster than both
lyrical music and instrumental music. The typing tests unfortuntely did not record which typing test of the five that each participant did so we
were unable to compare each test against another participant with the same test.

\section{Conclusion and Future Works}
% Don't forget to add results to the abstract as we cannot finsih until everything is done for the rest of the paper.
While the typing test was not statistically significant, there were at least two factors that should be changed should someone else recreate this study or 
further expand. The typing test should record which typing test a participant does. While the ones included are similar in length and difficulty, some 
participants said that some were easier becuase they recognized the lyrics from the songs used in the typing test and that they found the ones they 
recognized easier and faster for them to type. Although time may be a constraint, more text should be added to the typing test. It is not necessarily 
known if recognizable songs were a factor but older songs or different text completely could be implemented.

>>>>

During the cognitive task component of the study, participants performed a Quantified Behavioral test or QBTest while their eye movements were tracked by the EyeWriter 2.0.
The QBTest is a computerized test designed to assess various aspects of cognitive function, including attention, impulsivity, and response control.
The user will then see Figure~\ref{QBTest}. Figure~\ref{QBTest}a is the screen that will be displayed before the participant is ready to begin the QBTest.
Figure~\ref{QBTest}b is what the user will see during the test. Figure~\ref{QBTest}b also displays what the user will see when they are correct and incorrect.
Figure~\ref{QBTest}c displays for only a few seconds but shows the user their score.
The Quantified Behavioral test (QBTest) employed in this study is a computerized assessment tool designed to evaluate multiple facets of cognitive function, including attention, impulsivity, and response control. Developed specifically for this research, the QBTest presents participants with a series of tasks and stimuli designed to elicit cognitive responses while their eye movements are simultaneously tracked by the EyeWriter 2.0.

During the cognitive task component of the study, participants engage with the QBTest while their eye movements are monitored in real time. Figure~\ref{QBTest} illustrates the interface and feedback provided during the QBTest. Figure~\ref{QBTest}a depicts the initial screen displayed to participants before commencing the test, indicating readiness to begin. Once initiated, Figure~\ref{QBTest}b illustrates the interface presented to participants during the test, providing visual cues and prompts for task completion. Additionally, feedback for correct and incorrect responses is displayed to the user, facilitating real-time performance monitoring.

Furthermore, Figure~\ref{QBTest}c showcases the summary screen presented to participants after the two-minute testing period, displaying their overall score. This feedback mechanism enables participants to track their performance and provides researchers with quantitative data regarding cognitive function and task performance.

The QBTest serves as a valuable component of this study, allowing for assessing cognitive abilities in conjunction with eye-tracking data, thereby offering insights into the interplay between cognitive processes and ocular behaviors in individuals utilizing assistive technologies such as the EyeWriter 2.0.

\section{Results and Discussion}
Within-subjects testing was used for our ten participants. Individually, the typing test and QBTest scores were not statistically significant. However, the ANOVA conducted on the heatmap data obtained with the eyetracker revealed statistically significant results.

Interestingly, we observed that eye darting occurred most frequently in conditions with no music and music with lyrics. In contrast, the heatmap generated during music without lyrics showed a higher concentration of eye movements. This suggests that different auditory stimuli may influence participants' eye movements differently.

The increased occurrence of eye darting in conditions without music could indicate heightened visual scanning behavior, possibly due to reduced auditory distraction. Conversely, in conditions with music and lyrics, participants might experience a cognitive load associated with processing both auditory and visual information, leading to more frequent eye darting as they attempt to manage multiple stimuli.

On the other hand, the higher concentration observed in the heatmap for music without lyrics could imply focused attention or reduced visual distraction in response to the auditory stimulus. This finding aligns with research suggesting that instrumental music, devoid of lyrical content, may enhance concentration and cognitive performance by providing a soothing yet stimulating background.

These observations underscore the complex interplay between auditory stimuli and visual attention, highlighting the need for further investigation into how different types of music impact eye movements and cognitive processes.

@@

While not statistically significant, the \textit{p} value for time was 0.1172 whereas the \textit{p} value for accuracy was 0.2334.
The bar graph in Figure~\ref{timeL} for the time in seconds for the typing test was quite telling. The condition for no music was consistently faster than both
lyrical music and instrumental music. The typing tests unfortunately did not record which typing test of the five that each participant did so we
were unable to compare each test against another participant with the same test.

\section{Conclusion and Future Works}
In conclusion, our study investigated the influence of auditory stimuli on eye movements using an eyetracking device. While the typing test results were not statistically significant, our analysis of the heatmap data revealed intriguing patterns in participants' eye movements under different auditory conditions. Specifically, we observed that eye darting occurred most frequently in conditions with no music and music with lyrics, while music without lyrics led to a higher concentration of eye movements.

Looking ahead, there are several avenues for future research and improvements. Firstly, it would be beneficial to incorporate a recording of the specific typing test completed by each participant, as we suspect that familiarity with certain song lyrics may have influenced typing performance. Additionally, expanding the length and diversity of text in the typing test could provide further insights into how different stimuli affect typing accuracy and speed. Experimenting with a broader range of auditory stimuli, including older songs or entirely different texts, could help elucidate the factors influencing participants' eye movements.

Furthermore, exploring the cognitive mechanisms underlying the observed eye movement patterns and their relationship with auditory processing could enhance our understanding of multimodal perception and attention. Investigating individual differences in response to auditory stimuli, such as musical preferences or language proficiency, could also offer valuable insights into the variability of eye movement behavior across diverse populations.

In summary, our findings contribute to the growing body of research on the interplay between auditory stimuli and visual attention, highlighting the need for continued exploration and refinement of experimental methodologies to uncover the underlying mechanisms. By addressing these avenues for future research, we can further advance our understanding of human perception and enhance the design of technologies aimed at optimizing cognitive performance in various contexts.
